Inefficient Concessions and Mediation

### 2021-02-02 draft.tex
standardize math punctuation throughout appendix

If the potential advantages of ``Peace'' are large enough to outweigh the risks of encountering a low type in expected utility terms, that is $\de_h \geq \de^{nc}$, a county has the incentive to play Trust in the first round rather than Fight. Any player with $\de < \de_{nc}$ has the incentive---by this same inequality---to play Fight in each period even if the other country plays Trust. Thus we call these countries low types for purposes of this equilibrium. \hfill $\blacksquare$
%\noindent{\emph{Proof of Corollary \ref{corollary:1}}:\\
%High-type welfare under the no-concessions separating equilibrium is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D) \right\} = \frac{1}{1-\de_h}\left\{pT + (1-p)\left[W-D + \de_h(W-D)\right]\right\}$$
%High-type welfare under the grim trigger repeated game equilibrium is $\frac{1}{1-\de_h}\left\{pT + (1-p)\left[-D + \de_h(W-D) \right]\right\}$
%When $e=0$, welfare is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$$
%This type of separating equilibrium can exist for some sets of parameters. Since the high types have sufficient incentive to separate under appropriate parameter values even without concessions being given in period 0, we refer to such an equilibrium as a \emph{no-concessions separating equilibrium}. This is why the threshold patience level for this type of equilibrium is labeled $\de_h^{nc}$.

@@

High type separating equilibrium utility is $U_h= pX_{TT}+(1-p)X_{FF} - g_h+pg_h$. Since utility is decreasing in $g_h$, it is optimal for high types to make $g_h$ as small as possible and still have separation. This occurs when Expression~\ref{eq:cse} holds with equality. Substituting in the definitions of $X_{FT}^l$ and $X_{FF}^l$ yields the minimum separating gift:\footnote {While not formally addressed in this paper, inference about sizes of concessions when parties are of unequal force can be made. Note that the constraints determining the requisite size of the separating gift depend on the low types' ability to gather war spoils. Thus, a more powerful country would be more able to plunder, and hence has to make a greater concession to convincingly convey its type as being high. }
\be
	g_h^*=p(T+D) 
	\label{eq:cse2}
\ee 
 In order to incentivize high types to send nonzero concessions, the high type incentive compatibility constraint must not be violated. That is, the payoff from separating must be higher than the payoff from pooling with the low types. 
\be
	pX_{TT}^h+(1-p)X_{FF}^h - g_h +pg_h \geq X_{FF}^h +pg_h
	\label{eq:cse3}
\ee

$$p\frac{T}{1-\de_h}+(1-p)\frac{W-D}{1-\de_h} - g_h \geq \frac{W-D}{1-\de_h}$$
\begin{multline*}
	\frac{1}{1-\de_h}\left\{pT(1+g^1) + (1-p)(W-D(1+g^1))\right\} - g^1 \geq \\
	\frac{1}{1-\de_h}\left[p(W+Wg^1-D) + (1-p)(W-D) \right]
\end{multline*}
Combining Expressions~\ref{eq:cse2} and \ref{eq:cse3}, we have the condition for a concessions separating equilibrium to exist:
\be
pX_{TT}^h-pX_{FF}^h - p(X_{FT}^l-X_{FF}^l) \geq 0
\ee
In terms of fundamentals, the condition for a concessions separating equilibrium to exist is:
\be
%p_1(\frac{T+D-W}{1-\delta_h}-p_2(T+D)\geq 0. point is the p's arent the same; i might have gotten them backwards. simplified this is:
\delta_h \geq \frac{W}{T+D} = \de^c
\ee
That is, there is no parameter restriction beyond our definition of what it means for a player to be a high type. \hfill $\blacksquare$\\
\\
\emph{Proof of Corollary~\ref{corollary:1}:}\\
To show that the patience threshold for the \emph{no concessions separating equilibrium} is greater than the patience threshold for the \emph{concessions separating equilibrium}, we start with the assumption
$$T+D>W.$$
We can multiply both sides by $(1-p)>0$ and then subtract $p(T+D)$ from both sides to get
$$(T+D) > (1-p)W +p(T+D).$$
Since $0<p<1$ and W,T, and D are all positive, both sides of the inequality are positive and so we can divide through by both and multiply both sides by $W$:
$$\de^{nc} = \frac{W}{(1-p)W +p(T+D)} > \frac{W}{T+D} = \de^{c} $$

@@

High type separating equilibrium utility is $U_h= pX_{TT}+(1-p)X_{FF} - g_h+pg_h$. Since utility is decreasing in $g_h$, it is optimal for high types to make $g_h$ as small as possible and still have separation. This occurs when Expression~\ref{eq:matval2} holds with equality. That is, the optimal high type gift in this equilibrium is $g^1 \geq \frac{\left(1-\de_l\right)p\left(D +	 T\right)}{\left(1-\de_l\right)\left(1 - pT \right) + (1-p) D}$.\footnote {While not formally addressed in this paper, inference about sizes of concessions when parties are of unequal force can be made. Note that the constraints determining the requisite size of the separating gift depend on the low types' ability to gather war spoils. Thus, a more powerful country would be more able to plunder, and hence has to make a greater concession to convincingly convey its type as being high. }

In order to incentivize high types to send nonzero concessions, the high type incentive compatibility constraint must not be violated. That is, the payoff from separating must be higher than the payoff from not giving a gift and pooling with the low types. Note that if the high type pools with the low type, it knows it will be playing Fight and so invests any concession received in military buildup. 
\begin{multline*}
	\frac{1}{1-\de_h}\left\{pT(1+g^1) + (1-p)(W-D(1+g^1))\right\} - (1-p)g^1 \geq \\
	\frac{1}{1-\de_h}\left[p(W+Wg^1-D) + (1-p)(W-D) \right] +pg^1
\label{eq:matval3}
\end{multline*}

Simplifying, we have
$$pT(1+g^1) + (1-p)(W-D(1+g^1)) - (1-\de_h)g^1 \geq p(W+Wg^1-D) + (1-p)(W-D)$$
$$pT(1+g^1) - (1-p)Dg^1 - (1-\de_h)g^1 \geq p(W+Wg^1-D)$$
$$\de_h g^1 \geq p(W+Wg^1-D) - pT(1+g^1) + (1-p)Dg^1 + g^1$$
%$$\de_h \geq \frac{p(W+Wg^1-D) - pT(1+g^1) + (1-p)Dg^1 + g^1 }{g^1}$$
$$\de_h \geq \frac{p(W-D-T) + pg^1 (W-T) + (1-p)Dg^1 + g^1 }{g^1}$$
$$\de_h \geq \frac{p(W-D-T)}{g^1} + p (W-T) + (1-p)D + 1$$

@@

$$\frac{1}{1-\de_h}\left\{pT(1+eg_h) + (1-p)(W-D(1+eg_h))\right\} - (1-p)g_h$$

@@

For the general form of the low type incentive compatibility constraint, we start from Expression~\ref{eq:med} and add back in the $e$'s that were removed when assuming $e=1$ in the proof of Theorem~\ref{theorem:med}. So we start from
\begin{equation*}
			W-D \geq 
			\left(1-\de_l\right) \left[T + Teg_h + W + Weg_h \right] + \de_l \left(W + Weg_h -D \right) + peg_h - pg_h
		\end{equation*}
Simplifying, we have
\[
	pg_h\left(1 - e\right) \geq \left(T - \de_l T \right) \left(1 + eg_h\right) + Weg_h + \left(1 - \de_l \right) D
\]
Gathering the terms involving $g_h$ on the left hand side:
\[
	g_h\left[p\left(1 - e\right) -(1-\de_l)eT -eW \right] \geq \left(T - \de_l T \right) + \left(1 - \de_l \right) D
\]
The right hand side is positive. If the left hand side were negative, we would have a negative upper bound on the size of the gift. Given the requirement that the gift be non-negative, the low type's incentive compatibility constraint cannot be satisfied  in this case. Therefore we must assume that $p > e\left(p+\left(1-\de_l\right)T + W\right)$. Using this to isolate $g_h$ on the left hand side and gathering terms, we have 
\begin{equation}
	g_h \geq \frac{\left(1 - \de_l \right)\left(T +D \right)}{p - e\left(p +(1-\de_l)T -W \right)} 
	\label{eq:med3}
\end{equation}
If Condition~\ref{eq:med3} is met, the mechanism is incentive compatible for the low type.\\
(b) We now turn to the high type, whose incentive compatibility constraint is
\begin{equation}
	pX_{TT}^h+(1-p)X_{FF}^h - pg_h +peg_h \geq X_{FF}^h 
	\label{eq:med4}
\end{equation}
It is always optimal for the high type to invest in civil society when they truthfully reveal. When a high type misrepresents itself as a low type, the mechanism specifies that gifts are exchanged only when there are two high types. This means that there will be no concessions when a high type lies. This implies that Expression~\ref{eq:med4} is also the individual rationality constraint since outside the mechanism they get the `War' payoff and exchange no concessions.

>>>>

If the potential advantages of ``Peace'' are large enough to outweigh the risks of encountering a low type in expected utility terms, that is $\de_h \geq \de^{nc}$, a county has the incentive to play Trust in the first round rather than Fight. Any player with $\de < \de^{nc}$ has the incentive---by this same inequality---to play Fight in each period even if the other country plays Trust. Thus we call these countries low types for purposes of this equilibrium. \hfill $\blacksquare$
%\noindent{\emph{Proof of Corollary \ref{corollary:1}}:\\
%High-type welfare under the no-concessions separating equilibrium is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D) \right\} = \frac{1}{1-\de_h}\left\{pT + (1-p)\left[W-D + \de_h(W-D)\right]\right\}$$
%High-type welfare under the grim trigger repeated game equilibrium is $\frac{1}{1-\de_h}\left\{pT + (1-p)\left[-D + \de_h(W-D) \right]\right\}$
%When $e=0$, welfare is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$$
%This type of separating equilibrium can exist for some sets of parameters. Since the high types have sufficient incentive to separate under appropriate parameter values even without concessions being given in period 0, we refer to such an equilibrium as a \emph{no-concessions separating equilibrium}. This is why the threshold patience level for this type of equilibrium is labeled $\de_h^{nc}$.

@@

High type separating equilibrium utility is $U_h= pX_{TT}+(1-p)X_{FF} - g_h+pg_h$. Since utility is decreasing in $g_h$, it is optimal for high types to make $g_h$ as small as possible and still have separation. This occurs when Expression~\ref{eq:cse} holds with equality. Substituting in the definitions of $X_{FT}^l$ and $X_{FF}^l$ yields the minimum separating gift:\footnote {While not formally addressed in this paper, inference about sizes of concessions when parties are of unequal force can be made. Note that the constraints determining the requisite size of the separating gift depend on the low types' ability to gather war spoils. Thus, a more powerful country would be more able to plunder, and hence has to make a greater concession to convincingly convey its type as being high. }
\be
	g_h^*=p(T+D) 
	\label{eq:cse2}
\ee 
 In order to incentivize high types to send nonzero concessions, the high type incentive compatibility constraint must not be violated. That is, the payoff from separating must be higher than the payoff from pooling with the low types. 
$$pX_{TT}^h+(1-p)X_{FF}^h - g_h +pg_h \geq X_{FF}^h +pg_h$$
In terms of fundamentals, we have
\be
	p\frac{T}{1-\de_h}+(1-p)\frac{W-D}{1-\de_h} - g_h \geq \frac{W-D}{1-\de_h}
	\label{eq:cse3}
\ee
Combining Expressions~\ref{eq:cse2} and \ref{eq:cse3}, we have the condition for a concessions separating equilibrium to exist:
\be
%p_1(\frac{T+D-W}{1-\delta_h}-p_2(T+D)\geq 0. point is the p's arent the same; i might have gotten them backwards. simplified this is:
\delta_h \geq \frac{W}{T+D} = \de^c
\ee
That is, we simply need the high types to have a sufficiently large discount factor. \hfill $\blacksquare$\\
\\
\emph{Proof of Corollary~\ref{corollary:1}:}\\
To show that the patience threshold for the \emph{no concessions separating equilibrium} is greater than the patience threshold for the \emph{concessions separating equilibrium}, we start with the assumption $T+D>W.$

We multiply both sides by $(1-p)>0$ and then subtract $p(T+D)$ from both sides to get
$$(T+D) > (1-p)W +p(T+D).$$
Since $0<p<1$ and W,T, and D are all positive, both sides of the inequality are positive and so we can divide through by both and multiply both sides by $W$:
$$\de^{nc} = \frac{W}{(1-p)W +p(T+D)} > \frac{W}{T+D} = \de^{c} $$

@@

High type separating equilibrium utility is $U_h= pX_{TT}+(1-p)X_{FF} - g_h+pg_h$. Since utility is decreasing in $g_h$, it is optimal for high types to make $g_h$ as small as possible and still have separation. This occurs when Expression~\ref{eq:matval2} holds with equality. That is, the optimal high type gift in this equilibrium is $g^1 = \frac{\left(1-\de_l\right)p\left(D +	 T\right)}{\left(1-\de_l\right)\left(1 - pT \right) + (1-p) D}$.\footnote {While not formally addressed in this paper, inference about sizes of concessions when parties are of unequal force can be made. Note that the constraints determining the requisite size of the separating gift depend on the low types' ability to gather war spoils. Thus, a more powerful country would be more able to plunder, and hence has to make a greater concession to convincingly convey its type as being high. }

In order to incentivize high types to send nonzero concessions, the high type incentive compatibility constraint must not be violated. That is, the payoff from separating must be higher than the payoff from not giving a gift and pooling with the low types. Note that if the high type pools with the low type, it knows it will be playing Fight and so invests any concession received in military buildup. 
\begin{equation*}
	p\frac{T+Tg^1}{1-\de_h} + (1-p)\frac{W-D-Dg^1}{1-\de_h} - (1-p)g^1 \geq 
	p\frac{W+Wg^1-D}{1-\de_h} + (1-p)\frac{W-D}{1-\de_h} +pg^1
\label{eq:matval3}
\end{equation*}

Simplifying, we have
$$p(T+Tg^1) + (1-p)(W-D-Dg^1)) - (1-\de_h)g^1 \geq p(W+Wg^1-D) + (1-p)(W-D)$$
%$$pT(1+g^1) - (1-p)Dg^1 - (1-\de_h)g^1 \geq p(W+Wg^1-D)$$
$$\de_h g^1 \geq p(W+Wg^1-D) - pT(1+g^1) + (1-p)Dg^1 + g^1$$
%$$\de_h \geq \frac{p(W+Wg^1-D) - pT(1+g^1) + (1-p)Dg^1 + g^1 }{g^1}$$
$$\de_h \geq \frac{p(W-D-T) + pg^1 (W-T) + (1-p)Dg^1 + g^1 }{g^1}$$
$$\de_h \geq \frac{p(W-D-T)}{g^1} + p (W-T) + (1-p)D + 1$$

@@

$$p\frac{T+Teg_h}{1-\de_h} + (1-p)\frac{W-D-Deg_h}{1-\de_h} - (1-p)g_h$$

@@

For the general form of the low type incentive compatibility constraint, we start from Expression~\ref{eq:med} and add back in the $e$'s that were removed when assuming $e=1$ in the proof of Theorem~\ref{theorem:med}. So we start from
\begin{equation*}
			W-D \geq 
			\left(1-\de_l\right) \left[T + Teg_h + W + Weg_h \right] + \de_l \left(W + Weg_h -D \right) + peg_h - pg_h
		\end{equation*}
Simplifying, we have
\[
	pg_h\left(1 - e\right) \geq \left(T - \de_l T \right) \left(1 + eg_h\right) + Weg_h + \left(1 - \de_l \right) D
\]
Gathering the terms involving $g_h$ on the left hand side:
\[
	g_h\left[p\left(1 - e\right) -(1-\de_l)eT -eW \right] \geq \left(T - \de_l T \right) + \left(1 - \de_l \right) D
\]
The right hand side is positive. If the left hand side were negative, we would have a negative upper bound on the size of the gift. Given the requirement that the gift be non-negative, the low type's incentive compatibility constraint cannot be satisfied when the left hand side is negative and we must have that $p > e\left(p+\left(1-\de_l\right)T + W\right)$. Using this to isolate $g_h$ on the left hand side and gathering terms, we have 
\begin{equation}
	g_h \geq \frac{\left(1 - \de_l \right)\left(T +D \right)}{p - e\left(p +(1-\de_l)T -W \right)} 
	\label{eq:med3}
\end{equation}
If Condition~\ref{eq:med3} is met, the mechanism is incentive compatible for the low type.\\
(b) We now turn to the high type, whose incentive compatibility constraint is
\begin{equation}
	pX_{TT}^h+(1-p)X_{FF}^h - pg_h +peg_h \geq X_{FF}^h 
	\label{eq:med4}
\end{equation}
It is always optimal for the high type to invest in civil society when they truthfully reveal. When a high type misrepresents itself as a low type, the mechanism specifies that gifts are exchanged only when there are two high types. This means that there will be no concessions when a high type lies. This implies that Expression~\ref{eq:med4} is also the individual rationality constraint since outside the mechanism they get the `War' payoff and exchange no concessions.
